Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-06-18 parts-ccl.tex
remove mod commands

AANs are also equivalent to Logical Networks, that is,
either multivalued or Boolean networks
with evolution functions or focal parameters,
such as Thomas' models with Snoussi parameters.
This especially allows to efficiently compute reachability results
on large biological models,
provided that they are equivalent to Logical Networks,
which ensures that a translation to AANs is possible.
For example,
such a translation for generalized Interaction Graphs,
that is, Discrete Networks without evolution functions or parameters,
was proposed in~\cite{PMR10-TCSB}.

\modmf{%
The static analysis method proposed in this paper can theoretically be applied to other
kinds of formalisms that are not mentioned in the rest of this paper.
For example, AANs are expressive enough to represent Asynchronous Finite Cellular Automata,
that is, randomly updated and thus completely asynchronous Finite Cellular Automata.
Another possibility is to represent multiplayer games, which is possible
in AANs with 3 classes of priorities, with the addition of several automata
to tackle the rules of the game, and especially the sequentiality of the turns of each player.
In both cases, the considered formalism can be represented as an AAN,
making the application of our analysis method possible in theory.
However, in practice, the conclusiveness on such “unusual” models
can be potentially lowered, as the static analysis is not specifically tailored for them.
Indeed, some particular structures may create loops or non-independent nodes
in the Local Causality Graph,
making the whole method inconclusive.
However, more study and experiments are required to evaluate
the effectiveness of our method on these kinds of models.
}

Further work can be directly derived to improve the method presented in this paper.
The over-approximation on Process Hitting models without priorities proposed in~\cite{PMR12-MSCS}
and that was used in this work
is still accurate on AANs (by over-approximating dynamics)
but may be refined given the particular for of AANs proposed in this paper.
A specific search of key processes or cut sets \cite{PAK13-CAV} may especially be derived.
%but turns out to be too wide even in some obvious cases that are consequently not conclusive.
%This approximation may be refined in order to better fit the introduction of priorities, and mak the overall approximation approach 
%and mode precisely the class of models studied in this paper.
Furthermore, we are investigating alternative under-approximations that can be
applied directly to the whole class of AANs or Process Hitting models with priorities,
and not only to a sub-class with particular restrictions;
such improvement may permit to increase the conclusiveness of the static analysis
while allowing to analyse any model without the need of a translation.
Finally, in order to take into account quantitative data in transition delays, the overall approximation method could be extended to handle evolutions that are chronometric instead of only chronologic.
This may require the addition of information such as time delays in the model,
that would be exploited during the solving.

>>>>

AANs are also equivalent to Logical Networks, that is,
either multivalued or Boolean networks
with evolution functions or focal parameters,
such as Thomas' models with Snoussi parameters.
This especially allows to efficiently compute reachability results
on large biological models,
provided that they are equivalent to Logical Networks,
which ensures that a translation to AANs is possible.
For example,
such a translation for generalized Interaction Graphs,
that is, Discrete Networks without evolution functions or parameters,
was proposed in~\cite{PMR10-TCSB}.
AANs can also be used to represent sub-sets of
cellular automata \cite{cornforth_artificial_2003}
or multiplayer games \cite{gradel_infinite_2002},
although more experiments would be required to evaluate
the conclusiveness of our static analysis method on models of such particular forms.

Further work can also be directly derived to improve the method presented in this paper.
The over-approximation on Process Hitting models without priorities proposed in~\cite{PMR12-MSCS}
and that was used in this work
is still accurate on AANs (by over-approximating dynamics)
but may be refined given the particular for of AANs proposed in this paper.
A specific search of key local states or cut sets \cite{PAK13-CAV} may especially be derived.
%but turns out to be too wide even in some obvious cases that are consequently not conclusive.
%This approximation may be refined in order to better fit the introduction of priorities, and mak the overall approximation approach 
%and mode precisely the class of models studied in this paper.
Furthermore, we are investigating alternative under-approximations that can be
applied directly to the whole class of AANs or Process Hitting models with priorities,
and not only to a sub-class with particular restrictions;
such improvement may permit to increase the conclusiveness of the static analysis
while allowing to analyse any model without the need of a translation.
Finally, in order to take into account quantitative data in transition delays, the overall approximation method could be extended to handle evolutions that are chronometric instead of only chronologic.
This may require the addition of information such as time delays in the model,
that would be exploited during the solving.

### 2015-06-18 parts-sa.tex
remove mod commands

\modmf{%
The previous subsections \modLP{introduced the necessary definitions to build the
so-called} Local Causality Graph (LCG).
This graph is denoted $\cwB$ where $u$ is the local state to be reached
from the initial state $s$.
It will be the basis of our static analysis,
as it represents the causality links between the different objectives involved
in the solving of the reachability of $u$.
This LCG is built recursively by considering some required local states (\eg $u$),
linking them to objectives (\eg $\PHobj{t}{u}$, if $t \in s$),
and locally refining these objectives in order to include new required local states
from other automata.
An example of LCG is depicted in \pref{fig:sa-livelock}.
}

@@

\modmf{%
The definition of an LCG tackles two cases
where the target of an objective may be changed.
First,
the addition of objectives based on the local states already mentioned anywhere in the LCG,
as performed in \pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj},
ensures to take into account the possible changes in the active local states
made by other objectives.
We thus try to ensure that a required local state is reachable even when
starting from a local state of the same automaton that is not in the initial state,
but that may be active at some point of the solving.
Second, \pref{def:lcg}\nobreakdash-\ref{lcg-conn},
allows to “re-target” objectives whose own solving changes their target.
This happens when playing the actions that activate the local states required to solve
an objective $P$ also changes the active local state of automaton $\PHsort(P)$.
In this case, the initial objective $P$ is re-targeted to another
objective $\PHobj{p}{\bounce{P}}$, where $p$ is the new active local state in $\PHsort(P)$.
Linking to all objectives of this kind ensures that all possible disturbances
are taken into account.
}

\modmf{%
Finally, we note that
the LCG of \pref{def:lcg} contains synchronizing nodes ($\sSol$)
that allow to express the need for several local states simultaneously
in order to play a given action.
This is the main difference regarding~\cite{PMR12-MSCS,FPMR13-CS2Bio}
which tackled with Process Hitting, whose actions are limited to at most one hitter
and thus did not require this kind of synchronization.
}
%However, \pref{def:lcg} is equivalent to the definition of the structure
%given in~\cite{FPMR13-CS2Bio},
%although it was adapted for AANs which notably allowed to simplify its formalism.

>>>>

The previous subsections introduced the necessary definitions to build the
so-called Local Causality Graph (LCG).
This graph is denoted $\cwB$ where $u$ is the local state to be reached
from the initial state $s$.
It will be the basis of our static analysis,
as it represents the causality links between the different objectives involved
in the solving of the reachability of $u$.
This LCG is built recursively by considering some required local states (\eg $u$),
linking them to objectives (\eg $\PHobj{t}{u}$, if $t \in s$),
and locally refining these objectives in order to include new required local states
from other automata.
An example of LCG is depicted in \pref{fig:sa-livelock}.

@@

The definition of an LCG tackles two cases
where the target of an objective may be changed.
First,
the addition of objectives based on the local states already mentioned anywhere in the LCG,
as performed in \pref{def:lcg}\nobreakdash-\ref{lcg-ls-obj},
ensures to take into account the possible changes in the active local states
made by other objectives.
We thus try to ensure that a required local state is reachable even when
starting from a local state of the same automaton that is not in the initial state,
but that may be active at some point of the solving.
Second, \pref{def:lcg}\nobreakdash-\ref{lcg-conn},
allows to “re-target” objectives whose own solving changes their target.
This happens when playing the actions that activate the local states required to solve
an objective $P$ also changes the active local state of automaton $\PHsort(P)$.
In this case, the initial objective $P$ is re-targeted to another
objective $\PHobj{p}{\bounce{P}}$, where $p$ is the new active local state in $\PHsort(P)$.
Linking to all objectives of this kind ensures that all possible disturbances
are taken into account.

Finally, we note that
the LCG of \pref{def:lcg} contains synchronizing nodes ($\sSol$)
that allow to express the need for several local states simultaneously
in order to play a given action.
This is the main difference regarding~\cite{PMR12-MSCS,FPMR13-CS2Bio}
which tackled with Process Hitting, whose actions are limited to at most one hitter
and thus did not require this kind of synchronization.
